---
title: "final_eclipse_app"
output: html_document
date: "2023-03-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(skimr)
library(naniar)
library(here)
library(gtools)
library(RColorBrewer)
library(paletteer)
library(ggthemes)
library(shiny)
library(shinydashboard)
library(ggmap)
```


## Loading the Data

data website: https://www.kaggle.com/datasets/nasa/solar-eclipses

*Remember to change the path for loading the data between people*
```{r}
solar_eclipse <- read_csv("/Users/ksricket/Desktop/Astro_Project_BIS15L/NASA lunar and solar eclipses/solar.csv")
```
```{r}
lunar_eclipse <- read_csv(("/Users/ksricket/Desktop/Astro_Project_BIS15L/NASA lunar and solar eclipses/lunar.csv"), na =("-"))
```


## NAs
```{r}
naniar::miss_var_summary(solar_eclipse)
```
```{r}
naniar::miss_var_summary(lunar_eclipse)
```


## Variable Meanings:

#### Key to Catalog of Lunar Eclipses:
https://eclipse.gsfc.nasa.gov/LEcat5/LEcatkey.html

##### Key to Catalog of Solar Eclipses:
https://eclipse.gsfc.nasa.gov/SEcat5/SEcatkey.html



## Cleaning Names
```{r}
solar_e <- janitor::clean_names(solar_eclipse)
```
```{r}
names(solar_e)
```

```{r}
lunar_e <- janitor::clean_names(lunar_eclipse)
```
```{r}
names(lunar_e)
```




##Converting Dates: to date type data format, individual year, month, day variables
```{r}
solar_dates <- solar_e %>% 
  filter(!grepl("^-", calendar_date)) %>% 
  mutate(date = as.Date(calendar_date, format = "%Y %B %e")) %>% 
  select(!c(calendar_date)) %>% 
  mutate(year = lubridate::year(date),
         month = lubridate::month(date),
         day = lubridate::day(date))
solar_dates
```
```{r}
lunar_dates <- lunar_e %>% 
  filter(!grepl("^-", calendar_date)) %>% 
  mutate(date = as.Date(calendar_date, format = "%Y %B %e")) %>% 
  select(!c(calendar_date)) %>% 
  mutate(year = lubridate::year(date),
         month = lubridate::month(date),
         day = lubridate::day(date))
lunar_dates
```



## Converting Latitude and Longitude Coords
```{r}
lunar <- lunar_dates %>% 
  separate(latitude, into=c("lat", "NS"), sep="(?<=[0-9])(?=[A-Za-z])") %>% 
  separate(longitude, into=c("long", "EW"), sep="(?<=[0-9])(?=[A-Za-z])") %>% 
  mutate(lat_num = as.numeric(lat)) %>% 
  mutate(long_num = as.numeric(long)) %>% 
  mutate(latitude_value = ifelse(NS=="N", lat_num, -lat_num)) %>% 
  mutate(longitude_value = ifelse(EW=="E", long_num, -long_num)) 
lunar %>%
    select(latitude_value, longitude_value)
```
```{r}
solar <- solar_dates %>% 
  separate(latitude, into=c("lat", "NS"), sep="(?<=[0-9])(?=[A-Za-z])") %>% 
  separate(longitude, into=c("long", "EW"), sep="(?<=[0-9])(?=[A-Za-z])") %>% 
  mutate(lat_num = as.numeric(lat)) %>% 
  mutate(long_num = as.numeric(long)) %>% 
  mutate(latitude_value = ifelse(NS=="N", lat_num, -lat_num)) %>% 
  mutate(longitude_value = ifelse(EW=="E", long_num, -long_num))
solar %>%
    select(latitude_value, longitude_value)
```








## How to create a new categorical variable using case_when

```{r}
homerange <- homerange %>% 
             mutate(mass_category = case_when(log10_mass <= 1.75 ~ "small",
                                              log10_mass > 1.75 & log10_mass <= 2.75 ~ "medium",
                                              log10_mass > 2.75 ~ "large"))
#creates a new categorical variable based off of values from an existing variable
```
