---
title: "eclipse_project"
output: html_document
date: "2023-03-02"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(skimr)
library(naniar)
library(here)
library(gtools)
library(RColorBrewer)
library(paletteer)
library(ggthemes)
library(shiny)
library(shinydashboard)
library(ggmap)
```

## Loading the Data
```{r}
getwd()
```

data website: https://www.kaggle.com/datasets/nasa/solar-eclipses

```{r}
solar_eclipse <- read_csv("/Users/ksricket/Desktop/Astro_Project_BIS15L/NASA lunar and solar eclipses/solar.csv")
```
```{r}
lunar_eclipse <- read_csv("/Users/ksricket/Desktop/Astro_Project_BIS15L/NASA lunar and solar eclipses/lunar.csv")
```


## Cleaning Names
```{r}
solar_e <- janitor::clean_names(solar_eclipse)
```
```{r}
names(solar_e)
```

```{r}
lunar_e <- janitor::clean_names(lunar_eclipse)
```
```{r}
names(lunar_e)
```



## Exploring
```{r}
glimpse(solar_e)
```

```{r}
glimpse(lunar_e)
```

## Variable Meanings:

#### Key to Catalog of Lunar Eclipses:
https://eclipse.gsfc.nasa.gov/LEcat5/LEcatkey.html

##### Key to Catalog of Solar Eclipses:
https://eclipse.gsfc.nasa.gov/SEcat5/SEcatkey.html



## Converting Latitude and Longitude Coords
```{r}
lunar <- lunar_e %>% 
  separate(latitude, into=c("lat", "NS"), sep="(?<=[0-9])(?=[A-Za-z])") %>% 
  separate(longitude, into=c("long", "EW"), sep="(?<=[0-9])(?=[A-Za-z])") %>% 
  mutate(lat_num = as.numeric(lat)) %>% 
  mutate(long_num = as.numeric(long)) %>% 
  mutate(latitude_value = ifelse(NS=="N", lat_num, -lat_num)) %>% 
  mutate(longitude_value = ifelse(EW=="E", long_num, -long_num)) 
lunar %>%
    select(latitude_value, longitude_value)
```
```{r}
solar <- solar_e %>% 
  separate(latitude, into=c("lat", "NS"), sep="(?<=[0-9])(?=[A-Za-z])") %>% 
  separate(longitude, into=c("long", "EW"), sep="(?<=[0-9])(?=[A-Za-z])") %>% 
  mutate(lat_num = as.numeric(lat)) %>% 
  mutate(long_num = as.numeric(long)) %>% 
  mutate(latitude_value = ifelse(NS=="N", lat_num, -lat_num)) %>% 
  mutate(longitude_value = ifelse(EW=="E", long_num, -long_num))
solar %>%
    select(latitude_value, longitude_value)
```

## Convert Catalog Number to a Numeric 
(so that I can filter by catalog number later on as a test for mapping)
```{r}
solar$catalog_number_num <- as.numeric(solar$catalog_number)
```
```{r}
lunar$catalog_number_num <- as.numeric(lunar$catalog_number)
```


## Mapping Framework
```{r}
lunar %>% 
  select(latitude_value, longitude_value) %>% 
  summary()
```
```{r}
lat_l <- c(-70, 70)
long_l <- c(-180, 180)
bbox_l <- make_bbox(long_l, lat_l, f=0.05)
```
```{r}
map_l <- get_map(bbox_l, maptype = "terrain", source = "stamen")
```
```{r}
ggmap(map_l)
```


```{r}
solar %>% 
  select(latitude_value, longitude_value) %>% 
  summary()
```
```{r}
lat_s <- c(-80, 80)
long_s <- c(-180, 180)
bbox_s <- make_bbox(long_s, lat_s, f=0.05)
```
```{r}
map_s <- get_map(bbox_s, maptype = "toner-lite", source = "stamen")
```
```{r}
ggmap(map_s)
```

## Mapping Tests
```{r}
ggmap(map_s)+
  geom_point(data=solar, aes(longitude_value, latitude_value))
```
```{r}
solar_filtered_dates1 <- solar %>% 
  arrange(catalog_number_num) %>% 
  head(n=10)
```
```{r}
ggmap(map_s)+
  geom_point(data=solar_filtered_dates1, aes(longitude_value, latitude_value))
```
```{r}
solar_filtered_dates2 <- solar %>% 
  filter(catalog_number_num==1)
```
```{r}
ggmap(map_s)+
  geom_point(data=solar_filtered_dates2, aes(longitude_value, latitude_value))
```
```{r}
solar_filtered_dates3 <- solar %>% 
  filter(catalog_number_num<=10 & catalog_number_num>=1)
```
```{r}
ggmap(map_s)+
  geom_point(data=solar_filtered_dates3, aes(longitude_value, latitude_value, color=calendar_date, size=central_duration))+
  labs(x="Latitude",
       y="Longitude",
       title="Solar Eclipses 1-10",
       color="Date",
       size="Central Duration")
# we would need to filter out all observations with NA central duration from the very beginning - before even deciding year groupings
```
```{r}
names(solar)
```

https://www.rebeccabarter.com/blog/2017-11-17-ggplot2_tutorial/ 




## NA Summaries
```{r}
naniar::miss_var_summary(solar)
```
```{r}
naniar::miss_var_summary(lunar)
```



## Filtering by Eclipse Type
```{r}
lunar_filtered_type <- lunar %>% 
  filter(eclipse_type=="T" | eclipse_type=="T+" |eclipse_type=="T-")
lunar_filtered_type 
# only shows the total lunar eclipses
```
```{r}
lunar %>% 
  filter(grepl("P", eclipse_type))
# only shows partial lunar eclipses
```
```{r}
lunar %>% 
  filter(grepl("N", eclipse_type))
# only shows penumbral lunar eclipses
```


```{r}
solar_filtered_type <- solar %>% 
  filter(eclipse_type=="T" | eclipse_type=="Ts" |eclipse_type=="Tm" | eclipse_type=="Tn" | eclipse_type=="T-" | eclipse_type=="T+")
solar_filtered_type
# total solar eclipses
```
```{r}
solar %>% 
  filter(grepl("P", eclipse_type))
# partial solar eclipses
```
```{r}
solar %>% 
  filter(grepl("A", eclipse_type))
# annular solar eclipses
```
```{r}
solar %>% 
  filter(grepl("H", eclipse_type))
# hybrid (annular/total) solar eclipses
```






## Plots of Eclipse Type Proprtions
```{r}
solar %>% 
  ggplot(aes(x=eclipse_type))+
  geom_bar()
```

creating broad categories of eclipse types for more cohesive plots:
```{r}
solar %>% 
  mutate(eclipse_type_broad = case_when(grepl("A", eclipse_type) ~ "A",
                                        grepl("H", eclipse_type) ~ "H",
                                        grepl("P", eclipse_type) ~ "P",
                                        grepl("T", eclipse_type) ~ "T")) %>% 
  ggplot(aes(x=eclipse_type_broad, fill=eclipse_type_broad))+
  geom_bar()+
  labs(title="Proportions of Solar Eclipse Types",
       x="Solar Eclipse Type",
       y="Count",
       fill="Solar Eclipse Type")+
  scale_fill_discrete(name = "Solar Eclipse Type", labels = c("Annular", "Hybrid (Annular and Total)", "Partial", "Total"))
```



```{r}
lunar %>% 
  ggplot(aes(x=eclipse_type))+
  geom_bar()
```
```{r}
lunar %>% 
  mutate(eclipse_type_broad = case_when(grepl("N", eclipse_type) ~ "N",
                                        grepl("P", eclipse_type) ~ "P",
                                        grepl("T", eclipse_type) ~ "T")) %>% 
  ggplot(aes(x=eclipse_type_broad, fill=eclipse_type_broad))+
  geom_bar()+
  labs(title="Proportions of Lunar Eclipse Types",
       x="Lunar Eclipse Type",
       y="Count",
       fill="Lunar Eclipse Type")+
  scale_fill_discrete(name = "Lunar Eclipse Type", labels = c("Penumbral", "Partial", "Total"))
```




```{r}
?case_when()
```















---
title: "final_eclipse_app"
output: html_document
date: "2023-03-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(skimr)
library(naniar)
library(here)
library(gtools)
library(RColorBrewer)
library(paletteer)
library(ggthemes)
library(shiny)
library(shinydashboard)
library(ggmap)
```


## Loading the Data

data website: https://www.kaggle.com/datasets/nasa/solar-eclipses

*Remember to change the path for loading the data between people*
```{r}
getwd()
```
```{r}
solar_eclipse <- read_csv("/Users/katiericketts/Desktop/Astro_Project_BIS15L/NASA lunar and solar eclipses/solar.csv") %>% 
  clean_names()
```
```{r}
lunar_eclipse <- read_csv(("/Users/katiericketts/Desktop/Astro_Project_BIS15L/NASA lunar and solar eclipses/lunar.csv"), na =("-")) %>% 
  clean_names()
```
```{r}
solar_eclipse <- read_csv(here("NASA lunar and solar eclipses", "solar.csv")) %>% 
  clean_names()
```
```{r}
lunar_eclipse <- read_csv(here("NASA lunar and solar eclipses", "lunar.csv"), na=("-")) %>% 
  clean_names()
```


## NAs
```{r}
naniar::miss_var_summary(solar_eclipse)
```
```{r}
naniar::miss_var_summary(lunar_eclipse)
```


## Variable Meanings:

#### Key to Catalog of Lunar Eclipses:
https://eclipse.gsfc.nasa.gov/LEcat5/LEcatkey.html

##### Key to Catalog of Solar Eclipses:
https://eclipse.gsfc.nasa.gov/SEcat5/SEcatkey.html




##Converting Dates: to date type data format, individual year, month, day variables
```{r}
solar_dates <- solar_eclipse %>% 
  filter(!grepl("^-", calendar_date)) %>% 
  mutate(date = as.Date(calendar_date, format = "%Y %B %e")) %>% 
  select(!c(calendar_date)) %>% 
  mutate(year = lubridate::year(date),
         month = lubridate::month(date),
         day = lubridate::day(date))
solar_dates
```
```{r}
lunar_dates <- lunar_eclipse %>% 
  filter(!grepl("^-", calendar_date)) %>% 
  mutate(date = as.Date(calendar_date, format = "%Y %B %e")) %>% 
  select(!c(calendar_date)) %>% 
  mutate(year = lubridate::year(date),
         month = lubridate::month(date),
         day = lubridate::day(date))
lunar_dates
```



## Converting Latitude and Longitude Coords
```{r}
lunar <- lunar_dates %>% 
  separate(latitude, into=c("lat", "NS"), sep="(?<=[0-9])(?=[A-Za-z])") %>% 
  separate(longitude, into=c("long", "EW"), sep="(?<=[0-9])(?=[A-Za-z])") %>% 
  mutate(lat_num = as.numeric(lat)) %>% 
  mutate(long_num = as.numeric(long)) %>% 
  mutate(latitude_value = ifelse(NS=="N", lat_num, -lat_num)) %>% 
  mutate(longitude_value = ifelse(EW=="E", long_num, -long_num)) 
lunar %>%
    select(latitude_value, longitude_value)
```
```{r}
solar <- solar_dates %>% 
  separate(latitude, into=c("lat", "NS"), sep="(?<=[0-9])(?=[A-Za-z])") %>% 
  separate(longitude, into=c("long", "EW"), sep="(?<=[0-9])(?=[A-Za-z])") %>% 
  mutate(lat_num = as.numeric(lat)) %>% 
  mutate(long_num = as.numeric(long)) %>% 
  mutate(latitude_value = ifelse(NS=="N", lat_num, -lat_num)) %>% 
  mutate(longitude_value = ifelse(EW=="E", long_num, -long_num))
solar %>%
    select(latitude_value, longitude_value)
```



## Making the map
```{r}
latz <- c(-77, 77)
longz <- c(-164, 164)
bboxz <- make_bbox(longz, latz, f=0.05)
mapz <- get_map(bboxz, maptype = "toner-lite", source = "stamen")
```
```{r}
ggmap(mapz)
```
```{r}
solar %>% 
  
```



```{r}
lat <- c(-80, 80)
long <- c(-180, 180)
bbox <- make_bbox(long, lat, f=0.05)
map <- get_map(bbox, maptype = "toner-lite", source = "stamen")
```
```{r}
ggmap(map)
```


## Bare bones Dashboard shiny app

```{r}
library(shiny)

ui <- dashboardPage(
  dashboardHeader(title = "Peak Viewing Locations for Eclipses"),
  dashboardSidebar(title="Eclipse Selections", 
                   textInput("x", "Year", placeholder = "Type in a Year from 0-3000"),
                   selectInput("y", "Solar or Lunar Eclipses:", choices=c("Solar", "Lunar"), selected = "Solar"),
                   conditionalPanel(condition = "input.y == 'Solar'", radioButtons(
                     "z", label = "Which Type of Solar Eclipse", choices = c("Total", "Partial", "Annular", "Hybrid", "All Types"), selected = "All Types")
                    ),
                   conditionalPanel(condition = "input.y == 'Lunar'", radioButtons(
                     "a", label="Which Type of Lunar Eclipse", choices = c("Total", "Penumbral", "Partial", "All Types"), selected = "All Types")
                   )
                   ),
  dashboardBody(plotOutput("map", width = "600px", height = "700px"))
  
)

server <- function(input, output, session) {
  output$map <- renderPlot({
    app_lunar <- lunar %>%
      # NEW CODE
      mutate(year_char = as.character(year)) %>% 
      mutate(month_char = as.character(month)) %>%
      mutate(day_char = as.character(day)) %>%
      unite(date_char, "month_char", "day_char", "year_char", sep="/") %>% 
      filter(year == input$x)
    app_solar <- solar %>%
      # NEW CODE
      mutate(year_char = as.character(year)) %>% 
      mutate(month_char = as.character(month)) %>%
      mutate(day_char = as.character(day)) %>%
      unite(date_char, "month_char", "day_char", "year_char", sep="/") %>%
      filter(year==input$x)
    # Defining Solar Eclipse Sub-types 
    annular_solar <- app_solar %>% filter(grepl("A", eclipse_type))
    partial_solar <- app_solar %>% filter(grepl("P", eclipse_type))
    # NEW CODE
    total_solar <- app_solar %>% filter(grepl("T", eclipse_type))
    hybrid_solar <- app_solar %>% filter(grepl("H", eclipse_type))
    # Defining Lunar Eclipse Sub-types
    # NEW CODE
    total_lunar <- app_lunar %>% filter(grepl("T", eclipse_type))
    penumbral_lunar <- app_lunar %>% filter(grepl("N", eclipse_type))
    partial_lunar <- app_lunar %>% filter(grepl("P", eclipse_type))
    if (input$y == "Lunar") {
      if (input$a == "Total") {
        ggmap(map) + 
        geom_point(data=total_lunar, aes(longitude_value, latitude_value, color=date_char), size=4) +
        # NEW CODE
        labs(x= "Longitude", y= "Latitude", title="Total Lunar Eclipse Locations", color="Date (month/day/year)")
      }
      else if (input$a == "Partial") {
        ggmap(map) + 
        geom_point(data=partial_lunar, aes(longitude_value, latitude_value, color=date_char), size=4) +
        labs(x= "Longitude", y= "Latitude", title="Partial Lunar Eclipse Locations", color="Date (month/day/year)")
      }
      else if (input$a == "Penumbral") {
        ggmap(map) + 
        geom_point(data=penumbral_lunar, aes(longitude_value, latitude_value, color=date_char), size=4) +
        labs(x= "Longitude", y= "Latitude", title="Penumbral Lunar Eclipse Locations", color="Date (month/day/year)")
      } else {
        ggmap(map) + 
        geom_point(data=app_lunar, aes(longitude_value, latitude_value, color=date_char), size=4) +
        labs(x= "Longitude", y= "Latitude", title="All Lunar Eclipse Locations", color="Date (month/day/year)")
      }
    } else {
        if (input$z == "Annular") {
          ggmap(map) +
            geom_point(data=annular_solar, aes(longitude_value, latitude_value, color=date_char), size=4) +
            labs(x="Longitude", y="Latitude", title="Annular Solar Eclipse Locations", color="Date (month/day/year)")
        }
        else if (input$z == "Total") {
          ggmap(map) + 
          geom_point(data=total_solar, aes(longitude_value, latitude_value, color=date_char), size=4) +
          labs(x= "Longitude", y= "Latitude", title="Total Solar Eclipse Locations", color="Date (month/day/year)")
        }
        else if (input$z == "Partial") {
          ggmap(map) + 
          geom_point(data=partial_solar, aes(longitude_value, latitude_value, color=date_char), size=4) +
          labs(x= "Longitude", y= "Latitude", title="Partial Solar Eclipse Locations", color="Date (month/day/year)")
        }
        else if (input$z == "Hybrid") {
          ggmap(map) + 
          geom_point(data=hybrid_solar, aes(longitude_value, latitude_value, color=date_char), size=4) +
          labs(x= "Longitude", y= "Latitude", title="Hybrid Solar Eclipse Locations", color="Date (month/day/year)")
        } else {
          ggmap(map) + 
          geom_point(data=app_solar, aes(longitude_value, latitude_value, color=date_char), size=4) +
          labs(x= "Longitude", y= "Latitude", title="All Solar Eclipse Locations", color="Date (month/day/year)")
        }
      
    }
      })
    }
shinyApp(ui, server)
```

```{r}
solar %>% 
  mutate(date_factor =  as.factor(date)) %>% 
  select(date, date_factor)
```
```{r}
solar %>% 
  select(date, year, month, day)
```

```{r}
solar %>% 
  mutate(year_char = as.character(year)) %>% 
  mutate(month_char = as.character(month)) %>%
  mutate(day_char = as.character(day)) %>%
  unite(date_char, "month_char", "day_char", "year_char", sep="/") %>% 
  select(date_char)
```





# Other Data Analysis (For Presentation)

## Eclipse Types by Century

#### Creating Column Century (as numeric)
```{r}
solar_cent <- solar %>% 
  mutate(century_char = case_when(year >= 0 & year <100 ~ "1",
                                  year >= 100 & year <200 ~ "2",
                                  year >= 200 & year <300 ~ "3",
                                  year >= 300 & year <400 ~ "4",
                                  year >= 400 & year <500 ~ "5",
                                  year >= 500 & year <600 ~ "6",
                                  year >= 600 & year <700 ~ "7",
                                  year >= 700 & year <800 ~ "8",
                                  year >= 800 & year <900 ~ "9",
                                  year >= 900 & year <1000 ~ "10",
                                  year >= 1000 & year <1100 ~ "11",
                                  year >= 1100 & year <1200 ~ "12",
                                  year >= 1200 & year <1300 ~ "13",
                                  year >= 1300 & year <1400 ~ "14",
                                  year >= 1400 & year <1500 ~ "15",
                                  year >= 1500 & year <1600 ~ "16",
                                  year >= 1600 & year <1700 ~ "17",
                                  year >= 1700 & year <1800 ~ "18",
                                  year >= 1800 & year <1900 ~ "19",
                                  year >= 1900 & year <2000 ~ "20",
                                  year >= 2000 & year <2100 ~ "21",
                                  year >= 2100 & year <2200 ~ "22",
                                  year >= 2200 & year <2300 ~ "23",
                                  year >= 2300 & year <2400 ~ "24",
                                  year >= 2400 & year <2500 ~ "25",
                                  year >= 2500 & year <2600 ~ "26",
                                  year >= 2600 & year <2700 ~ "27",
                                  year >= 2700 & year <2800 ~ "28",
                                  year >= 2800 & year <2900 ~ "29",
                                  year >= 2900 & year <3000 ~ "30"))
solar_cent
```
```{r}
solar_cent$century = as.numeric(solar_cent$century_char)
```

#### Building the plot
```{r}
solar_cent %>% 
  ggplot(aes(x = century, fill = eclipse_type))+
  geom_bar()
```
#### Group the eclipse types by broader category
```{r}
solar_cent %>% 
  mutate(eclipse_type_broad = case_when(grepl("A", eclipse_type) ~ "A",
                                        grepl("H", eclipse_type) ~ "H",
                                        grepl("P", eclipse_type) ~ "P",
                                        grepl("T", eclipse_type) ~ "T")) %>%
  ggplot(aes(x = century, fill = eclipse_type_broad))+
  geom_bar()
```

## Proportions of Eclipse Types
```{r}
solar %>% 
  mutate(eclipse_type_broad = case_when(grepl("A", eclipse_type) ~ "A",
                                        grepl("H", eclipse_type) ~ "H",
                                        grepl("P", eclipse_type) ~ "P",
                                        grepl("T", eclipse_type) ~ "T")) %>% 
  ggplot(aes(x=eclipse_type_broad, fill=eclipse_type_broad))+
  geom_bar()+
  labs(title="Proportions of Solar Eclipse Types",
       x="Solar Eclipse Type",
       y="Count",
       fill="Solar Eclipse Type")+
  scale_fill_discrete(name = "Solar Eclipse Type", labels = c("Annular", "Hybrid (Annular and Total)", "Partial", "Total"))
```
```{r}
lunar %>% 
  mutate(eclipse_type_broad = case_when(grepl("N", eclipse_type) ~ "N",
                                        grepl("P", eclipse_type) ~ "P",
                                        grepl("T", eclipse_type) ~ "T")) %>% 
  ggplot(aes(x=eclipse_type_broad, fill=eclipse_type_broad))+
  geom_bar()+
  labs(title="Proportions of Lunar Eclipse Types",
       x="Lunar Eclipse Type",
       y="Count",
       fill="Lunar Eclipse Type")+
  scale_fill_discrete(name = "Lunar Eclipse Type", labels = c("Penumbral", "Partial", "Total"))
```









